Technical Report

### 2017-11-23 fully-abstract-comp/technical-report.tex
added a bit about the public transitions
Fix typo.
comment inserted
Resolved Lars' comment wrt K-relation
Added monotonicity requirements for memory interpretation functions in regions.
Proposed a definition for P in the LR
removed comments that we dealt with

% Function arrows
\newcommand{\fun}{\rightarrow}
\newcommand{\parfun}{\rightharpoonup}

@@

\newcommand{\future}{\sqsupseteq}
\newcommand{\pub}{\mathrm{pub}}
\newcommand{\privft}{\future^{\priv}}
\newcommand{\pubft}{\future^{\pub}}

@@

@@

\section{Logical Relation}
\subsection{Worlds}
\lau{30-10-2017: TODO: solve recursiveness in domain equation in the following. }
\lau{02-11-2017: We may need revoked regions again to handle the stack.}
\lau{06-11-2017: If we need to introduce revoked regions like we had for the local capabilities, we may consider the following: For the future world relation require the existence of a bijection from region names to region names. For all regions in the present world the future region is given by the bijection. This would mean that reinstantiation would happen simply by giving a bijection that maps to the new region. In other words, the weird shadowing would not be necessary.
  \begin{mathpar}
    \inferrule{ \dom(W') \supseteq \dom(W) \\ \exists m : \RegionName \fun \RegionName, \text{ $m$ injective} \ldotp \forall r \in \dom(W) \ldotp W'(m(r)) \pubft W(r) }
    { W' \pubft W }
  \end{mathpar}
}
\lau{10-11-2017: The above may not work well here because we do not want a $\spatialo$ region to be taken over by a $\pure$ region (as this would allow for normal capabilities addressing an area of memory that was only addressable by linear capabilities before).}
\[
  \World = \RegionName \parfun \Region
\]
where $\RegionName = \nats$.
\[
  \Region = \left\{
  \begin{array}{l}
    \{\spatial \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2}) \uplus \\
    \{\spatialo \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\uplus \\
    \{\pure \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2}) \uplus\\
    \{\revoked\}
  \end{array} \right.
\]
where $\spatial$ and $\spatialo$ are regions governing segments of memory addressed by linear capabilities. $\spatialo$ signifies that this region is addressable. $\spatial$ signifies that the region is not owned and can thus not be addressed. Finally, $\pure$ signifies that the region is only addressed by non-linear capabilities. Notice that no region allows for both linear and non-linear capabilities to address it.
\lau{26-20-2017: We may also want to build the infrastructure for the stack more explicitly into the worlds (but maybe we can handle it in the memory satisfaction).}
\lau{26-20-2017: It is not clear to me what the monotonicity requirements on the memory interpretation functions should be.}
  \[
   \Rels = \{\phi_\pub \times \phi \in \powerset{\State^2}\times \powerset{\State^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
  \]

@@

For both $\privft$ and $\pubft$ we have
\begin{mathpar}
  \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \future W(r) }
            { W' \future W }
\end{mathpar}

The private future regions allow transitions in the private part of the transition system. They also allow $\spatial$ regions to become $\revoked$:
\begin{mathpar}
  \inferrule{ }
            { \revoked \privft (\spatial,\_,\_,\_,\_)}
\and
  \inferrule{ }
            { \revoked \privft \revoked }
\and
  \inferrule{ (s,s') \in \phi }
            { (v,s',\phi^\pub,\phi,H) \privft (v,s,\phi^\pub,\phi,H)  }

\end{mathpar}

@@

\subsection{Relation}
\lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
\lau{19-10-2017: Maybe linearity should be taken care of in just one place (instead of in $\lrr$ relation and mem satisfaction). That is, add a predicate that checks whether the configuration satisfies linearity (such a condition may be to syntactic to work?).}
In the following definitions, \src{blue} is used to indicate values related to the source machine. This is unlike previous definitions, where \src{blue} is used  indicate source language specific parts of definitions.
\[
  \lre(W) = \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{k_{d,S}}, \src{k_{c,S}}, k_{d,T}, k_{c,T}, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrk, W_\lrr, W_\lrm \ldotp \\
      \qquad\npair[n']{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \in \lrp(W) \wedge \\
      \qquad\npair[n']{\stpair[.]{k_{d,S},k_{c,S}}{k_{d,T},k_{c,T}}} \in \lrk(W_\lrk) \wedge \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
      \qquad\Rightarrow\\
      \qquad\npair[n']{\left(\array{l}\src{(\ms_S,\reg_S\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,S},v_{d,S},k_{c,S},k_{d,S}},\stk, \ms_\stk)},\\
                                          (\ms_T,\reg_T\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,T},v_{d
,T},k_{c,T},k_{d,T}})\endarray\right)}
      \in \lro(W\oplus W_\lrk \oplus W_\lrr \oplus W_\lrm)
    \end{array}
    \right\}
\]
\lau{24-10-2017: Dominique, I do not see how the splitting of worlds is going to work out. I can see how it will enforce linearity, but is you give a "spatial" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfaction does not have it, and it cannot find a memory segment for this part. In other words, the memory satisfaction will not be able to find a concrete memory segment for segments governed by a linear capability in the register file. Maybe the memory satisfaction should have all of the parts, so it can represent them (but also limit what linear capabilities can be used).}

@@

\lau{03-11-2017: TODO take a look at linearity of this definition - in particular the recovered pc and stack pointer.}
\[
  \lrk(W) = \left\{ \npair{\left(\array{l}\src{\retptrd(\sigma,(\aaddr_d,\eaddr_d))},\\\src{\retptrc(\sigma,(\perm,\lin,\baddr_c,\eaddr_c,\aaddr_c))},\\k_{d,T},k_{c,T}\endarray\right)} \middle|
    \begin{array}{l}
      k_{d,T} = \sealed{\sigma,c_d} \wedge\\
      k_{c,T} = \sealed{\sigma,c_c} \wedge\\
      c_c = ((\perm,\lin),\baddr_c,\eaddr_c,\aaddr_c) \wedge\\
      c_d = ((\rw,\linear),\aaddr_d,\eaddr_d,\aaddr_d-1) \wedge\\
      \forall n' \leq n, W' \pubft W \ldotp \\
      \quad \forall W_\lrr, W_\lrm \ldotp  \\
      \qquad\opc = c_c \wedge\\
      \qquad\dom(\ms_{\stk,f}) = [\aaddr_d,\eaddr_d] \wedge\\
      \qquad\npair[n']{\src{\retptrd(\sigma,\_)},\src{\retptrc(\sigma,\_)},k_{d,T},k_{c,T}} \in \lrp(W') \wedge \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk::(\opc,\ms_{\stk,f}),\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
      \qquad\src{reg_S}(\rstk) = \src{\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\_}} \wedge \\
      \qquad\eaddr_\stk = \aaddr_d-1\\
      \qquad\Rightarrow\\
      \qquad\npair[n']{\left(\array{l}\src{\left(\arraycolsep=0pt\array{l}\ms_S,\reg_S\updReg{\pcreg,\rstk}{\opc,\stkptr{\rw,\baddr_\stk,\eaddr_d,\aaddr_d}},\\\stk, \ms_\stk \uplus \ms_{\stk,f}\endarray\right)},\\
                                          (\ms_T,\reg_T\updReg{\pcreg,\rstk}{\opc,((\rw,\lin),\baddr_\stk,\eaddr_d,\aaddr_d)})\endarray\right)}
      \in \lro(W' \oplus W_\lrr \oplus W_\lrm)
    \end{array}
  \right\}
\]

\[
  \lrp(W) = \left\{
    \npair{\stpair{v}{v}} \middle| \text{make sure linearity constraint satisfied}\dots 
  \right\}
\]

@@

\[
  \execCond{\lin,W} = \left\{ \npair{(S_\perm,\baddr,\eaddr)} \middle|
    \begin{array}{l}
      \forall n' < n, W' \privft W, \perm \in S_\perm, \aaddr \in [\baddr,\eaddr] \ldotp\\
      \quad \forall W'', c_{d,S},c_{d,T} \ldotp \npair[n']{(c_{d,S},c_{d,T})} \in \lrv(W'') \wedge \\
      \qquad \npair[n']{((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,S},((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,T}} \in \lrp(W' \oplus W'') \Rightarrow \\
      \quad \qquad \npair[n']{((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,S},((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,T}} \in \lre(W' \oplus W'')
    \end{array}
    \right\}
\]

>>>>

% Function arrows
\newcommand{\fun}{\rightarrow}
\newcommand{\parfun}{\rightharpoonup}
\newcommand{\monnefun}{\xrightarrow{\textit{\tiny{mon, ne}}}}

@@

\newcommand{\future}{\sqsupseteq}
\newcommand{\pub}{\mathrm{pub}}
\newcommand{\privft}{\future^{\priv}}
\newcommand{\pubft}{\future^{\pub}}
\newcommand{\monprivnefun}{\xrightarrow[\text{\tiny{$\privft$}}]{\textit{\tiny{mon, ne}}}}
\newcommand{\monpubnefun}{\xrightarrow[\text{\tiny{$\pubft$}}]{\textit{\tiny{mon, ne}}}}

@@

\[
  \pi_\baddr(\vsc) =
  \begin{cases}
    \baddr & \text{if $\vsc = ((\_,\_),\baddr,\_,\_)$} \\ 
    \pi_\baddr(c) & \text{if $\vsc = \sealed{\_,c}$} \\ 
    \src{\baddr} & \src{\text{if $\vsc = \stkptr{\_,\baddr,\_,\_}$}} \\ 
    \src{\baddr} & \src{\text{if $\vsc = \retptrd{\_,(\baddr,\_)}$}} \\ 
    \src{\baddr} & \src{\text{if $\vsc = \retptrc{\_,(\_,\_,\baddr,\_,\_)}$}}
  \end{cases}
\]
\[
  \pi_\eaddr(\vsc) =
  \begin{cases}
    \eaddr & \text{if $\vsc = ((\_,\_),\_,\eaddr,\_)$} \\ 
    \pi_\eaddr(c) & \text{if $\vsc = \sealed{\_,c}$} \\ 
    \src{\eaddr} & \src{\text{if $\vsc = \stkptr{\_,\_,\eaddr,\_}$}} \\ 
    \src{\eaddr} & \src{\text{if $\vsc = \retptrd{\_,(\_,\eaddr)}$}} \\ 
    \src{\eaddr} & \src{\text{if $\vsc = \retptrc{\_,(\_,\_,\_,\eaddr,\_)}$}}
  \end{cases}
\]
\[
  \pi_\lin(\vsc) = 
  \begin{cases}
    \lin & \text{if $\vsc = ((\_,\_),\_,\eaddr,\_)$} \\ 
    \pi_\lin(c) & \text{if $\vsc = \sealed{\_,c}$} \\ 
    \src{\linear} & \src{\text{if $\vsc = \stkptr{\_,\_,\eaddr,\_}$}} \\ 
    \src{\linear} & \src{\text{if $\vsc = \retptrd{\_,(\_,\_)}$}} \\ 
    \src{\lin} & \src{\text{if $\vsc = \retptrc{\_,(\_,\lin,\_,\_,\_)}$}}
  \end{cases}
\]

@@

\section{Logical Relation}
\subsection{Worlds}
\[
  \World = \RegionName \parfun \Region
\]
where $\RegionName = \nats$.
\[
  \Region = \left\{
  \begin{array}{l}
    \{\spatial \} \times \State \times \Rels \times (\State \fun \World \monprivnefun \UPred{\MemSeg^2}) \uplus \\
    \{\spatialo \} \times \State \times \Rels \times (\State \fun \World \monprivnefun \UPred{\MemSeg^2})\uplus \\
    \{\pure \} \times \State \times \Rels \times (\State \fun \World \monprivnefun \UPred{\MemSeg^2}) \uplus\\
    \{\revoked\}
  \end{array} \right.
\]
where $\spatial$ and $\spatialo$ are regions governing segments of memory addressed by linear capabilities. $\spatialo$ signifies that this region is addressable. $\spatial$ signifies that the region is not owned and can thus not be addressed. Finally, $\pure$ signifies that the region is only addressed by non-linear capabilities. Notice that no region allows for both linear and non-linear capabilities to address it.
\lau{26-20-2017: We may also want to build the infrastructure for the stack more explicitly into the worlds (but maybe we can handle it in the memory satisfaction).}
  \[
   \Rels = \{\phi_\pub \times \phi \in \powerset{\State^2}\times \powerset{\State^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
  \]

@@

For both $\privft$ and $\pubft$ we have
\begin{mathpar}
  \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \future W(r) }
            { W' \future W }
\end{mathpar}
%
The private future regions allow transitions in the private part of the transition system. They also allow $\spatial$ regions to become $\revoked$:
\begin{mathpar}
  \inferrule{ }
  { \revoked \privft (\spatial,\_,\_,\_,\_)}
\and
  \inferrule{ }
            { \revoked \privft \revoked }
\and
  \inferrule{ (s,s') \in \phi }
            { (v,s',\phi^\pub,\phi,H) \privft (v,s,\phi^\pub,\phi,H)  }
\end{mathpar}
%
Public future regions are allowed to transition according to the public part of their transition system.
\begin{mathpar}
    \inferrule{ }
            { \revoked \pubft \revoked }
\and
  \inferrule{ (s,s') \in \phi^\pub }
            { (v,s',\phi^\pub,\phi,H) \pubft (v,s,\phi^\pub,\phi,H)  }
\end{mathpar}

@@

\subsection{Relation}
\lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
\lau{19-10-2017: Maybe linearity should be taken care of in just one place (instead of in $\lrr$ relation and mem satisfaction). That is, add a predicate that checks whether the configuration satisfies linearity (such a condition may be to syntactic to work?).}
In the following definitions, \src{blue} is used to indicate values related to the source machine. This is unlike previous definitions, where \src{blue} is used  indicate source language specific parts of definitions.
\[
  \lre(W) = \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{k_{d,S}}, \src{k_{c,S}}, k_{d,T}, k_{c,T}, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrk, W_\lrr, W_\lrm \ldotp \\
      \qquad\npair[n']{\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}} \in \lrp(W) \wedge \\
      \qquad\npair[n']{\stpair[.]{k_{c,S},k_{d,S}}{k_{c,T},k_{d,T}}} \in \lrk(W_\lrk) \wedge \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
      \qquad\Rightarrow\\
      \qquad\npair[n']{\left(\array{l}\src{(\ms_S,\reg_S\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,S},v_{d,S},k_{c,S},k_{d,S}},\stk, \ms_\stk)},\\
                                          (\ms_T,\reg_T\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,T},v_{d
,T},k_{c,T},k_{d,T}})\endarray\right)}
      \in \lro(W\oplus W_\lrk \oplus W_\lrr \oplus W_\lrm)
    \end{array}
    \right\}
\]
\lau{24-10-2017: Dominique, I do not see how the splitting of worlds is going to work out. I can see how it will enforce linearity, but is you give a "spatial" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfaction does not have it, and it cannot find a memory segment for this part. In other words, the memory satisfaction will not be able to find a concrete memory segment for segments governed by a linear capability in the register file. Maybe the memory satisfaction should have all of the parts, so it can represent them (but also limit what linear capabilities can be used).}

@@

\lau{03-11-2017: TODO take a look at linearity of this definition - in particular the recovered pc and stack pointer.}

\[
  \lrk(W) = \left\{ \npair{\left(\array{l}\src{\retptrc(\sigma,(\perm,\lin,\baddr_c,\eaddr_c,\aaddr_c))},\\\src{\retptrd(\sigma,(\aaddr_d,\eaddr_d))},\\k_{c,T},k_{d,T}\endarray\right)} \middle|
    \begin{array}{l}
      k_{d,T} = \sealed{\sigma,c_d} \wedge\\
      k_{c,T} = \sealed{\sigma,c_c} \wedge\\
      c_c = ((\perm,\lin),\baddr_c,\eaddr_c,\aaddr_c) \wedge\\
      c_d = ((\rw,\linear),\aaddr_d,\eaddr_d,\aaddr_d-1) \wedge\\
      \npair[n]{\stpair[.]{\retptrc(\sigma,\_)}{k_{c,T}},\stpair[.]{\retptrc{\retptrd(\sigma,\_)}}{k_{d,T}}} \in \lrp(W) \wedge \\
      \forall n' \leq n, W' \pubft W \ldotp \\
      \quad \forall W_\lrr, W_\lrm \ldotp  \\
      \qquad\opc = c_c \wedge\\
      \qquad\dom(\ms_{\stk,f}) = [\aaddr_d,\eaddr_d] \wedge\\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk::(\opc,\ms_{\stk,f}),\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
      \qquad\src{reg_S}(\rstk) = \src{\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\_}} \wedge \\
      \qquad\eaddr_\stk = \aaddr_d-1\\
      \qquad\Rightarrow\\
      \qquad\npair[n']{\left(\array{l}\src{\left(\arraycolsep=0pt\array{l}\ms_S,\reg_S\updReg{\pcreg,\rstk}{\opc,\stkptr{\rw,\baddr_\stk,\eaddr_d,\aaddr_d}},\\\stk, \ms_\stk \uplus \ms_{\stk,f}\endarray\right)},\\
                                          (\ms_T,\reg_T\updReg{\pcreg,\rstk}{\opc,((\rw,\lin),\baddr_\stk,\eaddr_d,\aaddr_d)})\endarray\right)}
      \in \lro(W' \oplus W_\lrr \oplus W_\lrm)
    \end{array}
  \right\}
\]

\[
  \lrp(W) = \left\{
    \npair{\stpair[.]{c_1}{c_1'}\stpair[.]{c_2}{c_2'}} \middle| 
    \begin{array}{l}
      \text{for $i \in \{1,2\}$} \\
      \src{\baddr_i} = \pi_\baddr(\src{c_i}) \wedge \src{\eaddr_i} = \pi_\eaddr(\src{c_i}) \wedge 
      \baddr_i' = \pi_\baddr(c_i') \wedge \eaddr_i' = \pi_\eaddr(c_i') \wedge \\
      \src{\lin_i} = \pi_\lin(\src{c_i}) = \pi_\lin(c_i') \\
      \quad \exists r \in \addressable{l,W} \ldotp\\
      \qquad \forall \npair{n}{\stpair{\ms}{\ms}} \in W(r).H \; W(t).s \xi^{-1}(W)  \ldotp \\
      \qquad \quad \dom(\src{\ms_S}) \subseteq [\src{\baddr_i},\src{\eaddr_i}] \wedge \dom(\ms_T) \subseteq [\baddr_i',\eaddr_i']      
    \end{array}
  \right\}
\]

@@

\[
  \execCond{\lin,W} = \left\{ \npair{(S_\perm,\baddr,\eaddr)} \middle|
    \begin{array}{l}
      \forall n' < n, W' \privft W, \perm \in S_\perm, \aaddr \in [\baddr,\eaddr] \ldotp\\
      \quad \forall W'', c_{d,S},c_{d,T} \ldotp \npair[n']{\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lrv(W'') \wedge \\
      \qquad \npair[n']{((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,S},((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,T}} \in \lrp(W' \oplus W'') \Rightarrow \\
      \quad \qquad \npair[n']{\stpair[.]{((\perm,lin),\baddr,\eaddr,\aaddr)}{((\perm,lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lre(W' \oplus W'')
    \end{array}
    \right\}
\]
